---
title: "Basic R Data Visualization"
author: "Jonathan Johnston"
date: '2015-02-19'
output: html_document
---

TODO: Add table of contents.

### Introduction

Let's do some fun and basic data visualization in R! There's lots of learning to do here, so I'm curious to see how far I can go with the basics in R.

### Toy Data

I'm going to need some data to play with:

```{r}
# Set a seed so pseudo-random number generation is consistent
set.seed(1410)
# matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE, dimnames = NULL)
# runif(n, min = 0, max = 1) [n random numbers from unif dist]
# 'dimnames' argument specifies row and column names respectively
# LETTERS is a built-in constant of uppercase letters (letters is lowercase)
y <- matrix(runif(30), ncol=3, dimnames=list(letters[1:10], LETTERS[1:3]))
# Show the basic structure of the R object (in this case a matrix)
str(y)
# Show the data structures and types that I'm using
class(y)
class(y[1,1])
```

It's important to note (to myself mainly) that I'm using a matrix here, NOT a data frame. A data frame holds many different types of data, whereas a matrix only holds one data type; in this case the matrix only holds numerics.

### Visualization with 'Plot'

Now I can take a look at various ways of visualizing this toy data, starting with the basic 'plot' function. The specifications for plot are:

```
plot(x, y, ...)

* x - coordinates of points in the plot
* y - y coordinates of points in the plot (optional if x is appropriate structure)
* type - type of plot
    * "p" for points,
    * "l" for lines,
    * "b" for both,
    * "c" for the lines part alone of "b",
    * "o" for both ‘overplotted’,
    * "h" for ‘histogram’ like (or ‘high-density’) vertical lines,
    * "s" for stair steps,
    * "S" for other steps (vertical then horizontal),
    * "n" for no plotting.
* main - overall title
* sub - subtitle
* xlab - x axis label
* ylab - y axis label
* asp - y/x aspect ratio
```

I can also use 'sub()' to set or query graphical parameters.

```{r}
plot(y[,1], y[,2], main="A Simple Plot")
plot(y[,1], y[,2], main="A Plot with Axes Labels", xlab="A", ylab="B")
par(bg="thistle")
plot(y[,1], y[,2], main="A Coloured Etch-a-Sketch", xlab="A", ylab="B", type="s")
```

### Works Referenced

TODO: Link these nicely into the text.

1. [UC Riverside: Overview of Graphics and Data Visualization in R](http://faculty.ucr.edu/~tgirke/HTML_Presentations/Manuals/Rgraphics/Rgraphics.pdf)
2. [Markdown Cheatsheet](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet)
3. [R Matrices](http://stat.ethz.ch/R-manual/R-devel/library/base/html/matrix.html)
4. [R Uniform Distribution](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/Uniform.html)
5. [Constants built into R](http://stat.ethz.ch/R-manual/R-devel/library/base/html/Constants.html)
6. [View R structure with 'str()'](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/str.html)
7. [Matrices versus Data Frames](https://stackoverflow.com/questions/5158790/data-frame-or-matrix)
8. [R Data Frames](http://stat.ethz.ch/R-manual/R-devel/library/base/html/data.frame.html)
9. [R Data Types](http://www.statmethods.net/input/datatypes.html)
10. [R Generic X-Y Plotting](http://stat.ethz.ch/R-manual/R-devel/library/graphics/html/plot.html)
11. [Set or Wuery Graphical Parameters in R](http://stat.ethz.ch/R-manual/R-devel/library/graphics/html/par.html)
12. [R Markdown Basics](http://rmarkdown.rstudio.com/authoring_basics.html)